Pyro-Velocity

### 2024-07-15 nbs/about/contributing.qmd
docs(contributing): improve notes regarding devshell usage
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

Using [nix](https://github.com/NixOS/nix), [home-manager](https://github.com/nix-community/home-manager), and [nixpkgs](https://github.com/NixOS/nixpkgs) [dockerTools](https://nixos.org/manual/nixpkgs/stable/#sec-pkgs-dockerTools), [nixpod](https://ghcr.io/cameronraysmith/nixpod) provides a containerized or containerizable drop-in user configuration on any platform where the [nix](https://github.com/NixOS/nix) package manager is already, or can be, [installed](https://nix.dev/install-nix.html). This is intended to include, but is not limited to, scenarios like those involving [kubernetes ephemeral containers](https://kubernetes.io/docs/concepts/workloads/pods/ephemeral-containers/) via images like [netshoot](https://github.com/nicolaka/netshoot), which might be used for debugging purposes adjacent to otherwise minimal container images. See the [github packages associated to this repository](https://github.com/cameronraysmith?tab=packages&repo_name=nixpod) for the containers currently built in the [nixpod CID](https://github.com/cameronraysmith/nixpod/blob/main/.github/workflows/cid.yaml).

This repository originally contained a [nix flake](https://zero-to-nix.com/concepts/flakes) that essentially integrated a few parts of [srid/nixos-config](https://github.com/srid/nixos-config) into [juspay/nix-dev-home](https://github.com/juspay/nix-dev-home). These were merged upstream in [juspay/nix-dev-home#7](https://github.com/juspay/nix-dev-home/pull/7), so you might want to look there.

@@

The instructions above only cover python dependencies.
We use the [nix](https://github.com/NixOS/nix) package manager with a [nix flake](https://nix.dev/concepts/flakes) to manage both system 
dependencies and python dependencies together in a reproducible manner.
The nix flake integrates with [direnv](https://github.com/direnv/direnv) and provides a devlopment shell that can be activated with

@@

if you have [nix](https://nix.dev/install-nix.html) and direnv installed.

@@

We provide a relatively minimal [OCI image](https://github.com/opencontainers/image-spec)
built with [nix](#nix) and [dockerTools](https://nixos.org/manual/nixpkgs/stable/#sec-pkgs-dockerTools).
It only supports linux x86-64, but can be run with rosetta on macOS.

>>>>

@@

If you do not know what [nix](https://nixos.org/) is and are interested in using this section, 
please see [this video for a brief introduction to nix flakes](https://youtu.be/JCeYq72Sko0?si=adRGX0YYm1Ejmd4N&t=96).

One of the reasons you might be interested is that the instructions above only cover python dependencies.
We use the [nix](https://github.com/NixOS/nix) package manager with a [nix flake](https://nix.dev/concepts/flakes) to manage both system 
dependencies and python dependencies together in a reproducible manner.
This requires referencing the lock files in [flake.lock](https://github.com/pinellolab/pyrovelocity/blob/main/flake.lock)
and [poetry.lock](https://github.com/pinellolab/pyrovelocity/blob/main/poetry.lock) with the bridge provided by 
[poetry2nix](https://github.com/nix-community/poetry2nix).
At present there isn't really another python package management system with a sufficiently
detailed lock file to ensure that the python dependencies are reproducibly installable in concert with
the system-level dependencies. Please see the rejected [PEP 665](https://peps.python.org/pep-0665/) 
and other discussions that reference it as this may change in the near future.
Taken together, this is why we currently use poetry to manage python dependencies in development environments.

Before proceeding ensure you have deactivated any virtual 
environments created with python-ecosystem tools.
The nix flake integrates with [direnv](https://github.com/direnv/direnv) and provides a
[nix devlopment shell](https://github.com/pinellolab/pyrovelocity/blob/v0.2.0/flake.nix#L185) 
that can be activated with

@@

if you have [nix](https://nix.dev/install-nix.html) and [direnv](https://github.com/direnv/direnv/tree/v2.34.0#basic-installation) installed.
Please note this will require at least 10GB of free hard disk space to build and activate the environment.
However, if your disk is within 50GB of being full, you may encounter a slow build process.
Since you will need to enable the usage of nix flakes, you should either install
the nix package manager with the 
[Determinate Systems nix installer](https://github.com/DeterminateSystems/nix-installer/tree/v0.20.1?tab=readme-ov-file#installation-differences)
or update your `/etc/nix/nix.conf` or other active nix configuration to enable 
the `nix-command` and `flakes` features.

```nix
experimental-features = nix-command flakes
```

These features are not really "experimental" in any usual sense of the term
so you should not be concerned about enabling them.

@@

We provide a relatively minimal [OCI image](https://github.com/opencontainers/image-spec)
built with [nix](#nix) and [dockerTools](https://nixos.org/manual/nixpkgs/stable/#sec-pkgs-dockerTools).
It only supports linux x86-64, but can be run via rosetta emulation on macOS.
